A Note on GMM Estimation of Theories of Expectation Formation

### 2019-10-31 notes/GMMResults.tex
derived stochastic vol version of RE, SE, NI. and updated the GMM Est Notes.

\subsection{Different models of expectation formation}

\subsubsection{Rational expectation}

@@

\begin{eqnarray}
\begin{split}
\hat y^*_{t+h|i,t} & =  E^*_{i,t}(y_{t+h}|I_{i,t}) = \tau_t 
\end{split}
\end{eqnarray}

@@

\begin{eqnarray}
\begin{split}
Var^*_{t+h|i,t} & = \sum^{h}_{t=0} E_{i,t}(\sigma^2_{\eta,t+h}) +  E_{i,t}(\sigma^2_{\epsilon,t+h})  \\
& = \sum^{h}_{t=1} E_{i,t}(exp^{\log \sigma^2_{\eta,t}+\sum^h_{t=1}\mu_{\eta,t+h}}) +  E_{i,t}(exp^{\log \sigma^2_{\epsilon,t}+\sum^h_{t=0}\mu_{\epsilon,t+h}} ) \\
& = \sigma^2_{\eta,t} \sum^{h}_{t=0} E_{i,t}(exp^{\sum^h_{t=1}\mu_{t+h,\eta}}) +  \sigma^2_{\epsilon,t} E_{i,t}(exp^{\sum^h_{t=1}\mu_{\epsilon,t+h}} ) \\
& = \sigma^2_{\eta,t} \sum^{h}_{t=0} exp^{E_{i,t}({\sum^h_{t=1}\mu_{t+h,\eta}})- 0.5Var_{i,t}(\sum^h_{t=1}\mu_{t+h,\eta})} +  \sigma^2_{\epsilon,t} E_{i,t}(exp^{\sum^h_{t=1}\mu_{\epsilon,t+h}} ) \\
& = \sigma^2_{\eta,t} \sum^{h}_{t=0} exp^{- 0.5h\gamma_{\eta}} +  \sigma^2_{\epsilon,t} exp^{- 0.5h\gamma_{\epsilon}}  \\
\end{split} 
\end{eqnarray}

For instance, set $h=0$, the conditional volatility for the current inflation is

\begin{eqnarray}
\begin{split}
Var^*_{t|i,t} =  \sigma^2_{\eta,t}  +  \sigma^2_{\epsilon,t}
\end{split} 
\end{eqnarray}

>>>>

\subsection{Rational expectation}

@@

\begin{eqnarray}
	\begin{split}
		\overline y^*_{t+h|t} \equiv y^*_{t+h|i,t} & =  E^*_{i,t}(y_{t+h}|I_{i,t}) = \theta_t 
	\end{split}
\end{eqnarray}

Forecast error is simply the cumulated sum of unrealized shocks from $t$ to $t+h$, which is

\begin{eqnarray}
	\begin{split}
		\overline{FE}^*_{t+h|t} \equiv  FE^*_{t+h|i,t} & =  \sum^{h}_{s=1} (\eta_{t+s} + \epsilon_{t+s})
	\end{split}
\end{eqnarray}

@@

\begin{eqnarray}
	\begin{split}
		\overline{Var}^*_{t+h|t} \equiv  Var^*_{t+h|i,t} & = \sum^{h}_{k=1} E_{i,t}(\sigma^2_{\eta,t+k}) +  E_{i,t}(\sigma^2_{\epsilon,t+k})  \\
		& = \sum^{h}_{k=1} E_{i,t}(exp^{\log \sigma^2_{\eta,t}+\sum^h_{k=1}\mu_{\eta,t+k}}) +  E_{i,t}(exp^{\log \sigma^2_{\epsilon,t}+\sum^h_{f=1}\mu_{\epsilon,t+f}} ) \\
		& = \sum^{h}_{k=1}\sigma^2_{\eta,t} E_{i,t}(exp^{\sum^h_{k=1}\mu_{t+h,\eta}}) +  \sigma^2_{\epsilon,t} E_{i,t}(exp^{\sum^h_{f=1}\mu_{\epsilon,t+f}} ) \\
		& = \sum^{h}_{k=1}\sigma^2_{\eta,t}  exp^{E_{i,t}({\sum^h_{k=1}\mu_{t+k,\eta}})- 0.5Var_{i,t}(\sum^h_{k=1}\mu_{t+k,\eta})} +  \sigma^2_{\epsilon,t} E_{i,t}(exp^{\sum^h_{f=1}\mu_{\epsilon,t+f}} ) \\
		& = \sigma^2_{\eta,t} \sum^{h}_{k=1} exp^{- 0.5k\gamma_{\eta}} +  \sigma^2_{\epsilon,t} exp^{- 0.5h\gamma_{\epsilon}} 
	\end{split} 
\end{eqnarray}

One immediately see that now the volatility is stochastic at any point of the time.

For instance, set $h=1$, the conditional volatility for the 1-step-ahead inflation is

\begin{eqnarray}
	\begin{split}
		Var^*_{t+1|i,t} =  exp^{- 0.5\gamma_{\eta}} \sigma^2_{\eta,t}  +  exp^{- 0.5\gamma_{\epsilon}} \sigma^2_{\epsilon,t} 
	\end{split} 
\end{eqnarray}

Disgreement is zero across agents in RE.

\begin{eqnarray}
	\begin{split}
		\overline{Disg}^*_{t+h|t} =  0 
	\end{split} 
\end{eqnarray}

\subsection{Sticky Expectation}

An agent whose most recent up-do-date update happened at $t-\tau$, thus she sees all the realizations of stochastic variables up to $t-\tau$, including $y_{t-\tau}$, $\tau_{t-\tau}$, $\eta_{t-\tau}$, $\sigma_{\eta,t-\tau}$, $\sigma_{\epsilon,t-\tau}$.

Her forecast is the permanent component that realized at time $t-\tau$.

\begin{eqnarray}
\begin{split}
y_{t+h|i,t-\tau}  = \theta_{t-\tau} 
\end{split}
\end{eqnarray}

Her forecast uncertainty is

\begin{eqnarray}
\begin{split}
Var_{t+h|i,t-\tau} & = \sigma^2_{\eta,t-\tau} \sum^{h+\tau}_{k=1} exp^{- 0.5k\gamma_{\eta}} +  \sigma^2_{\epsilon,t-\tau} exp^{- 0.5(h+\tau)\gamma_{\epsilon}}
\end{split} 
\end{eqnarray}

The population average of the two are, respectively, a weighted average of people whose the most update was in $t, t-1... t-\tau, t-\infty$, respectively.

\begin{eqnarray}
\begin{split}
\overline y^{se}_{t+h|t} & = \sum^{\infty}_{\tau=0} (1-\lambda)^\tau\lambda y_{t+h|t-\tau} \\
& = \sum^{\infty}_{\tau=0} (1-\lambda)^\tau\lambda \theta_{t-\tau}
\end{split} 
\end{eqnarray}

\begin{eqnarray}
\begin{split}
\overline {Var}^{se}_{t+h|t} & = \sum^{\infty}_{\tau=0} (1-\lambda)^\tau\lambda Var_{t+h|t-\tau} \\
& = \sum^{\infty}_{\tau=0} (1-\lambda)^\tau\lambda [ \sigma^2_{\eta,t-\tau} \sum^{h+\tau}_{k=1} exp^{- 0.5k\gamma_{\eta}} +  \sigma^2_{\epsilon,t-\tau} exp^{- 0.5(h+\tau)\gamma_{\epsilon}}]
\end{split} 
\end{eqnarray}

Both forecast errors $\overline{FE}_{t+h|t}$ and disagreements takes similar form to that in AR process with time-invariant volatility.

\begin{eqnarray}
\begin{split}
\overline {FE}^{se}_{t+h|t} & = \sum^{\infty}_{\tau=0} (1-\lambda)^\tau\lambda {FE}^*_{t+h|t-\tau} 
& = \sum^{\infty}_{\tau=0} (1-\lambda)^\tau\lambda \sum^{\tau+h}_{s=1} (\theta_{t+s} + \epsilon_{t+s})  
\end{split} 
\end{eqnarray}

The disagreement is the following.

\begin{eqnarray}
\begin{split}
\overline{Disg}^{se}_{t+h|t} & =  \sum^{\infty}_{\tau=0} (1-\lambda)^{2\tau} \lambda^2 (y_{t+h|t-\tau} - \overline y^{se}_{t+h|t})^2  \\
& = \sum^{\infty}_{\tau=0} (1-\lambda)^{2\tau} \lambda^2 (\theta_{t-\tau} - \overline y^{se}_{t+h|t})^2  \\
& = \sum^{\infty}_{\tau=0} (1-\lambda)^{2\tau} \lambda^2 \{\theta_{t-\tau} - \sum^{\infty}_{\tau=0} (1-\lambda)^\tau\lambda \theta_{t-\tau}\}^2  
\end{split} 
\end{eqnarray}

\subsection{Noisy Information}

Now, the agent at time $t$ needs to recover the real-time permanent component $\theta_t$ to make the best forecast for future $y_{t+h}$ using nosiy signals.

\begin{eqnarray}
y^{ni}_{t+h|t}  \equiv  y^{ni}_{t+h|i,t} = \bar \theta_{t|t}
\end{eqnarray}

where $\bar \theta_{t|t}$ is generated through Kalman filtering.

Assume that the nosiy signals of $\theta_t$ consists of a public signals $s^{pb}_{t}$ and the private signals $s^{pr}_{i,t}$. For simplicity, let us assume the public signal is basically the $y_t$. A more general case would be an independently drawn public signal sequence. The two signals can be again stacked into a vector of $2\times 1$ to $s^\theta_{i,t}$.

Then the filterred $\theta_{t}$ by agent $i$ is

\begin{eqnarray}
\begin{split}
\bar \theta_{t|t} =  (1-\tilde P_{t} H) \bar \theta_{t|t-1} + \tilde P s^\theta_{i,t} \\
\end{split}
\end{eqnarray}

where $\theta_{t|t-k}$ is the filterred forecast of $\theta_t$ using all the information up to $t-k$., and $\tilde Pt$ is the time-specific Kalman gain that is dependent on the noisy ratios of signals.

\begin{eqnarray}
\begin{split}
\tilde P_{t} =  \Sigma^\theta_{i,t|t-1} H(H'\Sigma^\theta_{i,t|t-1} H + \Sigma^\theta_{t})^{-1} 
\end{split}
\end{eqnarray}

Now the noisiness of signals are time varying as well.

\begin{eqnarray}
\begin{split}
\Sigma^\theta_t =  \left[ \begin{matrix} 
\sigma^2_{\eta,t} &  0 \\ 
0 & \sigma^2_\xi \end{matrix}\right] 
\end{split}               
\end{eqnarray}

where the variance of public signal is the time-varying $\sigma^2_{\eta,t}$ and let us assume the private signals have constant nosiness $\sigma^2_{\xi}$.

The uncertainty matrix evolves in the following manner.

\begin{eqnarray}
\Sigma^\theta_{i,t|t} = \Sigma^\theta_{i,t|t-1} - \Sigma^\theta_{i,t|t-1} H'(H \Sigma^\theta_{i,t-1} H' +\Sigma^\theta_{t}) H \Sigma^\theta_{i,t|t-1} 
\end{eqnarray}

Notice now that since prior of $\theta_t$ before time period $t$ is a weighted average of previous realizations of $y$ and past signals, current forecast depends on the past realizations even though the rational forecast is up-to-date $\theta_t$.

Due to the time-varying volatility $\sigma^2_{\eta,t}$, the noisyness of the public signal is also time-varying, which governs the degree of rigidity. For instance, if volatility to the transitory shock is high, the Kalman gain is low, the forecast responds less to the new realizations.

It is also worth thinking about what is the one-step ahead uncertainty in the context of stochastic valotility world.

\begin{eqnarray}
\begin{split}
\Sigma^\theta_{i,t|t-1} & = \Sigma^\theta_{i,t-1|t-1} + Var^*_{t|t-1}(y_t) \\
& = \Sigma^\theta_{i,t-1|t-1} +  exp^{- 0.5\gamma_{\eta}} \sigma^2_{\eta,t-1}  +  exp^{- 0.5\gamma_{\epsilon}} \sigma^2_{\epsilon,t-1} 
\end{split}
\end{eqnarray}
